Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2014-01-20 ms.tex
lots of clean up and some re-organization

The \PLANCK~2013 data release \citep{planck2013} represents an important 
opportunity to revisit foreground predictions in light of \PLANCK's superb, 
relatively artifact-free broadband data covering the entire sky and a wide 
range of frequencies. \cite{planckdust} has recently conducted a study modeling
\PLANCK~353 GHz, 545 GHz, 857 GHz and IRAS 100$\mu$m emission with a single
MBB spectrum. Here we investigate the FDS99 two-component dust emission model 
as an alernative model for the dust spectral energy distribution (SED) composed
 of \PLANCK~and \IRAS~data. In doing so, we obtain $6.1'$ resolution maps of 
dust temperature and optical depth. Because we employ a model that has been 
validated with FIRAS data down to sub-mm wavelengths, we expect our best-fit 
parameters to be useful in constructing high-resolution predictions of dust 
emission over a very broad range of wavelengths. This includes low frequencies 
($<$350 GHz), which \cite{planckdust} caution their model may not adequately 
fit, and also measurements near the peak of the dust SED, for example 
\AKARI~140-160$\mu$m \citep{akari}. We also anticipate our derived optical 
depth map will serve as a useful comparison for extinction estimates based 
directly upon optical/NIR observations of stars.

@@

In $\S$\ref{sec:data} we discuss the data used throughout this study. In 
$\S$\ref{sec:prepro} we describe our preprocessing of the data to isolate 
thermal emission from Galactic dust. In $\S$\ref{sec:modeling} we describe the 
dust emission model we fit to the \PLANCK~and \IRAS~data. In 
$\S$\ref{sec:fitting} we provide details of the Markov Chain Monte Carlo method
with which we have estimated the parameters of our model. In 
$\S$\ref{sec:results} we present the full-sky maps of dust temperature and 
optical depth thus obtained, and conclude and $\S$\ref{sec:conclusion}

%how what we're doing fits in
%     where FDS99 fit in
%     Planck data qualities that improve on data used in FDS99
%     benefits of using Planck + FDS99 model
%utility/applications of what we're doing

%list of figures : 
% posterior surface for some pixel, with samples from markov chain overplotted
%    --> also overplot output best-fit (T,I545) and unctertainty estimate
% plot for some set of pixels of SED overlaid with best-fit two-component model
%    --> done in Python instead of IDL
%    --> overlay 
% plot of resulting T2 and tau
% plots of full-sky residuals relative to our model ?
%  ---> and for lower-frequency bands to emphasize that FDS99 is better for
%       extrapolating towards mm wavelengths
% zoom-in of region with interesting dust showing :
%  SFD T map, two-component T map, Planck team one-component T map
%      good spot is in polaris, centered on (lgal,bgal) ~ (112, 20)

\section{Data}
\label{sec:data}
All \PLANCK~data products utilized throughout this work are drawn from the 
\PLANCK~2013 data release \citep{planck2013}. Specifically, we have made use 
of the the 217 GHz, 353 GHz, 545 GHz and 857 GHz intensity maps. We use the 
versions of these maps which have been corrected for zodiacal light 
\citep[\texttt{R1.10\_nominal\_ZodiCorrected},][]{planckzodi}.

%TODO: actually smooth all of the maps to a common PSF of 6.1 arcmin.

To incorporate information about the dust emission on the Wien side of its 
spectrum, we make use of 100$\mu$m data. In particular, we use the SFD 
reprocessing of IRAS 100$\mu$m, which we will refer to as \verb|i100|, and at 
times by frequency as 3000 GHz. The \verb|i100| map has angular resolution of 
$6.1'$. 
%list other %important characteristics of i100 e.g. that it has been 
%zodi-corrected and relies on dirbe on scales larger than ???

\section{Pre-processing}
\label{sec:prepro}

@@

\subsection{CMB Removal}
The CMB is effectively imperceptible at 857 GHz, but can be noticed at a 
low level at 545 GHz, and is prominent at 217 GHz and 353 GHz relative 
to the diffuse Galactic emission we wish to characterize, especially at high
latitudes. To remove the CMB anisotropies, we have subtracted the SMICA model
from each of the relevant \PLANCK~maps \citep{smica}, after appropriate unit 
conversions. Removal of the residual Solar dipole at low frequencies is 
discussed in $\S$\ref{sec:zp}.

\subsection{Finite Bandpass Correction}
The thermal dust emission model of $\S$\ref{sec:modeling} predicts the 
intensity $M_{\nu}$ in MJy/sr for any single frequency $\nu$. But in reality
we are comparing to broad bands typically hundreds of GHz wide. The shape of 
model spectrum $M_{\nu}$ is only dependent upon one parameter: the hot
dust temperature $T_2$. We have therefore computed bandpass correction factors
$b_{\nu_i}$ such that $M_{\nu_i}b_{\nu_i}$ is the predicted observation for the
\PLANCK~or \IRAS~band of interest. We have used the \PLANCK~transmission 
curves provided in the file \verb|HFI_RIMO_R1.10.fits| \citep{planckresponse}. 
For 3000 GHz, we have used the \IRAS~bandpass, following the treatment of
\cite{planckdust}. We note that it is not clear which bandpass should
properly be considered for the IRAS/DIRBE composite 100$\mu$m maps used in 
this analysis (here SFD \verb|i100|) and that of \cite{planckdust}.

@@

With $i$ denoting a single nside=64 HEALPix pixel in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{217}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and constant offset removed, smoothed to $1^{\circ}$ resolution. 
$\mathcal{I}_{217}$ is the corresponding $1^{\circ}$ resolution \PLANCK~217 
GHz map with zodiacal emission and compact sources removed. $\mathcal{D}$ is a
scaling of the Solar dipole with unit amplitude. Because $\sim$18,000 pixels
satisfy the low HI cut, we have an overconstrained linear model with three
parameters: a, b, and c. $a$ represents the slope of \PLANCK~217 GHz 
versus \PLANCK~857 GHz assuming they are linearly correlated, and is not
of any importance to our downstream analysis. $b$ represents the scaling of 
the Solar dipole that makes the 857 GHz vs. 217 GHz correlation tightest, and 
$c$ represents the constant offset necessary to make the zero level of the 217 
GHz map consistent with that of the 857 GHz map. Performing a linear least
squares fit with uniform weights and iterative outlier rejection, we find 
$a=4.33\times10^{-5}\ K/(MJy/sr)$, $b=-1.40\times10^{-5}\ K$, and 
$c=8.12\times10^{-5}\ K$. Note that $b$ represents 0.42\% of the Solar dipole 
amplitude, which is within the calibration uncertainty of the \PLANCK~217 GHz 
band. Figure 1 shows scatterplots of the 857 GHz versus 217 GHz correlation 
before and after correcting for our best-fit offset and dipole residual. Not 
only is the improvement striking in this scatterplot, but the residual dipole 
also appears visually well-corrected in the 217 GHz map itself. Before 
performing thermal dust fits, we therefore subtract $c$ and $b\mathcal{D}$ from
the 217 GHz map.

@@

\begin{figure*}
\begin{center}
\epsfig{file=posteriors.eps, width=7.0in}
\caption{Gridded posterior PDFs for three nside=2048 pixels. Red crosses mark 
the best-fit parameters based on our Markov Chain sampling of the posterior. 
The posterior distributions are in general extremely well-behaved, showing
no multimodality or other pathological qualities. Our MCMC parameter 
estimates coincide well with the peaks in the gridded posteriors. The 
colorscale is linear in $log(P)$, with black representing the maximum of 
$log(P)$ and white representing $max(log(P))-5$. Left: Low S/N pixel at 
high-latitude in Galactic north. Center: High S/N pixel in the Polaris region. 
Right: Low S/N pixel at high-latitude in the Galactic south.}
\end{center}
\end{figure*}

@@

At each sky location, we have five independent intensity measurements (four ) 
FDS99 found typical values for $T_2$ and

@@

>>>>

re
The \PLANCK~2013 data release \citep{planck2013} represents an important 
opportunity to revisit foreground predictions in light of \PLANCK's superb, 
relatively artifact-free broadband data covering the entire sky and a wide 
range of frequencies. \cite{planckdust} has recently conducted a study modeling
\PLANCK~353 GHz, 545 GHz, 857 GHz and \IRAS~100$\mu$m emission with a single
MBB spectrum. Here we investigate the FDS99 two-component dust emission model 
as an alernative model for the dust spectral energy distribution (SED) composed
 of \PLANCK~and \IRAS~data. In doing so, we obtain $6.1'$ resolution maps of 
dust temperature and optical depth. Because we employ a model that has been 
validated with FIRAS data down to sub-mm wavelengths, we expect our best-fit 
parameters to be useful in constructing high-resolution predictions of dust 
emission over a very broad range of wavelengths. This includes low frequencies 
($<$350 GHz), which \cite{planckdust} caution their model may not adequately 
fit, and also measurements near the peak of the dust SED, for example 
\AKARI~140-160$\mu$m \citep{akari}. We also anticipate our derived optical 
depth map will serve as a useful comparison for extinction estimates based 
directly upon optical/NIR observations of stars.

@@

In $\S$\ref{sec:data} we discuss the data used throughout this study. In 
$\S$\ref{sec:prepro} we describe our preprocessing of the data to isolate 
thermal emission from Galactic dust. In $\S$\ref{sec:modeling} we describe the 
dust emission model we fit to the \PLANCK~and \IRAS~data. In 
$\S$\ref{sec:fitting} we provide details of the Markov chain Monte Carlo method
with which we have estimated the parameters of our model. In 
$\S$\ref{sec:results} we present the full-sky maps of dust temperature and 
optical depth thus obtained, and conclude in $\S$\ref{sec:conclusion}.

 \section{Data}
\label{sec:data}
All \PLANCK~data products utilized throughout this work are drawn from the 
\PLANCK~2013 release \citep{planck2013}. Specifically, we have made use 
of the 217 GHz, 353 GHz, 545 GHz and 857 GHz intensity maps which have been 
corrected for zodiacal light 
\citep[\texttt{R1.10\_nominal\_ZodiCorrected},][]{planckzodi}.

To incorporate measurements on the Wien side of the dust emission spectrum, 
we include 100$\mu$m data in our SED fits. In particular, we use the SFD 
reprocessing of \IRAS~100$\mu$m, which we will refer to as \verb|i100|, and at 
times by frequency as 3000 GHz. The \verb|i100| map has angular resolution of 
$6.1'$, and was constructed so as to contain only thermal 
emission from Galactic dust, with compact sources and zodiacal light removed, 
and its zero level tied to HI. We use the \verb|i100| map as-is, without any 
custom modifications.
%maybe clarify that throughout we use SFD i100 totally as-is, with no
% customization

%list other %important characteristics of i100 e.g. that it has been 
%zodi-corrected and relies on dirbe on scales larger than ???

\section{Preprocessing}
\label{sec:prepro}

@@

\subsection{CMB Removal}
The CMB is effectively imperceptible in \PLANCK~857 GHz, but can be noticed at 
a low level in \PLANCK~545 GHz, and is prominent in \PLANCK~217 GHz and 353 
GHz relative to the diffuse Galactic emission we wish to characterize, 
especially at high latitudes. To remove the CMB anisotropies, we have 
subtracted the SMICA model from each of the relevant \PLANCK~maps 
\citep{smica}, after appropriate unit conversions. Removal of the residual 
Solar dipole at low frequencies is discussed in $\S$\ref{sec:zp}.

@@

With $i$ denoting a single nside=64 HEALPix pixel in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{217}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and constant offset removed, smoothed to $1^{\circ}$ resolution. 
$\mathcal{I}_{217}$ is the corresponding $1^{\circ}$ resolution \PLANCK~217 
GHz map with zodiacal emission and compact sources removed. $\mathcal{D}$ is a
scaling of the Solar dipole with unit amplitude. Because $\sim$18,000 pixels
satisfy the low HI cut, we have an overconstrained linear model with three
parameters: a, b, and c. $a$ represents the slope of \PLANCK~217 GHz 
versus \PLANCK~857 GHz assuming they are linearly correlated, and is not
of any importance to our downstream analysis. $b$ represents the scaling of 
the Solar dipole that makes the 857 GHz vs. 217 GHz correlation tightest, and 
$c$ represents the constant offset necessary to make the zero level of the 217 
GHz map consistent with that of the 857 GHz map. Performing a linear least
squares fit with uniform weights and iterative outlier rejection, we find 
$a=4.33\times10^{-5}\ K/(MJy/sr)$, $b=-1.40\times10^{-5}\ K$, and 
$c=8.12\times10^{-5}\ K$. Note that $b$ represents 0.42\% of the Solar dipole 
amplitude, which is within the calibration uncertainty of the \PLANCK~217 GHz 
band. Figure \ref{fig:scatter} shows scatterplots of the 857 GHz versus 217 
GHz correlation before and after correcting for our best-fit offset and dipole 
residual. Not only is the improvement striking in this scatterplot, but the 
residual dipole also appears visually well-corrected in the 217 GHz map 
itself. Before performing thermal dust fits, we therefore subtract $c$ and 
$b\mathcal{D}$ from the 217 GHz map.

@@

\begin{figure*}
\begin{center}
\epsfig{file=posteriors.eps, width=7.0in}
\caption{Gridded posterior PDFs for three nside=2048 pixels. Red crosses mark 
the best-fit parameters based on our Markov chain sampling of the posterior. 
The posterior distributions are in general extremely well-behaved, showing
no multimodality or other pathological qualities. Our MCMC parameter 
estimates coincide well with the peaks in the gridded posteriors. The 
colorscale is linear in $log(P)$, with black representing the maximum of 
$log(P)$ and white representing $max(log(P))-5$. Left: Low S/N pixel at 
high-latitude in Galactic north. Center: High S/N pixel in the Polaris region. 
Right: Low S/N pixel at high-latitude in the Galactic south.}
\end{center}
\end{figure*}

@@

At each sky location, we have five independent intensity measurements (four ) 
FDS99 found typical values for $T_2$ and 
%basically just finish clarifying here that we want to start simple by fitting
%only two parameters for each pixel since we only have 5 intensity
% measurements per pixel in our baseline fits at 6.1 arcmin resoluiton

@@

\subsection{Finite Bandpass Correction}
The thermal dust emission model of $\S$\ref{sec:modeling} predicts the 
intensity $M_{\nu}$ in MJy/sr for any single frequency $\nu$. But in reality
we are comparing to broad bands typically hundreds of GHz wide. The shape of 
model spectrum $M_{\nu}$ is only dependent upon one parameter: the hot
dust temperature $T_2$. We have therefore computed bandpass correction factors
$b_{\nu_i}$ such that $M_{\nu_i}b_{\nu_i}$ is the predicted observation for the
\PLANCK~or \IRAS~band of interest. We have used the \PLANCK~transmission 
curves provided in the file \verb|HFI_RIMO_R1.10.fits| \citep{planckresponse}. 
For 3000 GHz, we have used the \IRAS~bandpass, following the treatment of
\cite{planckdust}. We note that it is not clear which bandpass should
properly be considered for the \IRAS/DIRBE composite 100$\mu$m maps used in 
this analysis (here SFD \verb|i100|) and that of \cite{planckdust}.
